Exploiting Private and Hybrid Clouds for Data-intensive Computing

### 2012-08-04 MSc-dissertation.tex
"August"

With more than 2.5 quintillion bytes of data \cite{website:ibm-bigdata} being produced
every day, every now and then we run into a problem of not having 
sufficient computing resources to process and extract useful information
within our required time-frame.

The term big-data is time specific and broadly
big data can be defined at any point in time as ``data whose size forces us
 to look beyond the tried-and-true methods
that are prevalent at that time.`` In the early 1980s, it was a dataset that was
 so large that a robotic ``tape monkey'' was required to swap thousands of tapes in and out.
In the 1990s, perhaps, it was any data that transcended the bounds of Microsoft
Excel and a desktop PC, requiring serious software on Unix workstations to
analyze. Nowadays, it may mean data 
that is too large to be placed in a relational database and analyzed with the
help of a desktop statistics/visualization package-data, perhaps, whose
analysis requires massively parallel software running on tens, hundreds, or
even thousands of servers \cite{jacobs2009pathologies}.

The term big-data is more relevant today because 90\% \cite{website:ibm-bigdata}
of the data in the world today has been created in the last two years and 
hence big-data is getting a lot of hype from business publications and IT blogs
and the mainstream media.

@@

According to IBM, the challenges of big-data spans three dimensions: Volume,
Velocity and Variety \cite{website:ibm-bigdata}.

@@

As early as 1978, the idea of using multiple autonomous computers to solve a problem were being
moulded. With the  advent  of  microcomputers and computer networks, distributed processing
became technically and economically feasible \cite{chu1978distributed}

@@

Virtualization technologies are 
used to subdivide the ample resources of a modern computer and present the illusion of many
smaller virtual machines(VMs)\cite{barham2003xen}.
Virtualization has been applied to operating systems both commercially and in research for nearly forty years.
IBM VM/370 first made use of virtualization to allow binary support for legacy code \cite{gum1983virt}.
Advances is virtualization technologies has enabled computing models like the cloud computing which uses
the virtual machine technology that partition a machine to support the concurrent
execution of multiple operating systems.

% Types of virtualization?
% ----------------------
There are mainly two types of virtualization: \cite{matthews2007quantifying}

@@

 \item \emph{Paravirtualization} \\
  In paravirtualization, specific changes are made in the interface provided to a virtual machine to avoid some features that are difficult or expensive to virtualize. This requires that modifications be made in the operating system to deal with this modified interface. The Xen hypervisor \cite{website:xen} uses paravirtualization.
\end{itemize}

@@

%Hardware assisted virtualization
Hardware assisted virtualization is becoming common and is used to get better efficency when doing virtualization. Hardware vendors are rapidly embracing virtualization and developing new features
to simplify virtualization techniques. First generation enhancements include Intel
Virtualization Technology (VT-x) and AMD's AMD-V which both target privileged 
instructions with a new CPU execution mode feature that allows the VMM to run in a privileged mode and thus  sensitive calls are set to automatically trap to the hypervisor, removing the need for either binary translation or paravirtualization \cite{website:vmware}

@@

Virtualization provides enables better IT infrastructure management with benefits like \cite{website:intel-VT}:

@@

Cloud Computing was a concept first publicly announced by John McCarthy in 1961 
(in a speech given to celebrate MIT's centennial). He saw a future in which computing 
power and even specific applications could be sold through the utility 
business model like water or electricity \cite{garfinkel1999architects}.
This concept and  supporting technologies evolved over the period and has taken us 
to the reality where computing resources are served as a utility. 
Technologies relating to CPU Power, network bandwidth, memory capacity 
and connectivity have improved dramatically in this period enabling Cloud Computing as 
a business model.

Cloud Computing is of interest among IT and network technologists
and academics, as well as CIOs and business leaders, due to its potential to 
reduce cost and risk, increase revenue, and enhance total customer experience.\cite{weinman2011future}
It has been difficult to give a concise definition for Cloud Computing 
and related terms. According to the U.S. National Institute of Standards 
and Technology (NIST), \cite{mell2009nist}  "Cloud computing is a model for enabling convenient,
on-demand network access to a shared pool of configurable computing resources
(e.g., networks, servers, storage, applications, and services) that can 
be rapidly provisioned and released with minimal management 
effort or service provider interaction."

@@

The essential characteristics are \cite{mell2009nist} :

@@

Apache Hadoop \cite{website:apache-hadoop} is a top-level Apache project that includes open source implementations of a distributed file system \cite{website:apache-hdfs} and MapReduce that was inspired by Google's GFS
\cite{ghemawat2003google} and MapReduce \cite{dean2008mapreduce} projects.

@@

Facebook uses the Apache Hadoop platform to run it messages application \footnote {Facebook Messages gives every 
Facebook user a facebook.com email address, integrates the display of all e-mail, SMS and chat messages between a pair or group of users, has strong controls over who users receive messages from, and is the foundation of a Social Inbox\cite{borthakur2011apache}.}. Hadoop was chosen because it met the application's reqirements for consistency, availability, partition tolerance, data model and scalability \cite{borthakur2011apache}.

@@

@@

>>>>

With more than 2.5 quintillion bytes of data\cite{website:ibm-bigdata} being produced
every day, every now and then we run into a problem of not having 
sufficient computing resources to process and extract useful information
within our required time-frame.

The term big-data is time specific and broadly
big data can be defined at any point in time as ``data whose size forces us
 to look beyond the tried-and-true methods
that are prevalent at that time.`` In the early 1980s, it was a dataset that was
 so large that a robotic ``tape monkey'' was required to swap thousands of tapes in and out.
In the 1990s, perhaps, it was any data that transcended the bounds of Microsoft
Excel and a desktop PC, requiring serious software on Unix workstations to
analyze. Nowadays, it may mean data 
that is too large to be placed in a relational database and analyzed with the
help of a desktop statistics/visualization package-data, perhaps, whose
analysis requires massively parallel software running on tens, hundreds, or
even thousands of servers\cite{jacobs2009pathologies}.

The term big-data is more relevant today because 90\%\cite{website:ibm-bigdata}
of the data in the world today has been created in the last two years and 
hence big-data is getting a lot of hype from business publications and IT blogs
and the mainstream media.

@@

According to IBM, the challenges of big-data spans three dimensions: Volume,
Velocity and Variety\cite{website:ibm-bigdata}.

@@

As early as 1978, the idea of using multiple autonomous computers to solve a problem were being
moulded. With the  advent  of  microcomputers and computer networks, distributed processing
became technically and economically feasible\cite{chu1978distributed}

@@

Virtualization technologies are 
used to subdivide the ample resources of a modern computer and present the illusion of many
smaller virtual machines(VMs)\cite{barham2003xen}.
Virtualization has been applied to operating systems both commercially and in research for nearly forty years.
IBM VM/370 first made use of virtualization to allow binary support for legacy code\cite{gum1983virt}.
Advances is virtualization technologies has enabled computing models like the cloud computing which uses
the virtual machine technology that partition a machine to support the concurrent
execution of multiple operating systems.

% Types of virtualization?
% ----------------------
There are mainly two types of virtualization:\cite{matthews2007quantifying}

@@

 \item \emph{Paravirtualization} \\
  In paravirtualization, specific changes are made in the interface provided to a virtual machine to avoid some features that are difficult or expensive to virtualize. This requires that modifications be made in the operating system to deal with this modified interface. The Xen hypervisor\cite{website:xen} uses paravirtualization.
\end{itemize}

@@

%Hardware assisted virtualization
Hardware assisted virtualization is becoming common and is used to get better efficency when doing virtualization. Hardware vendors are rapidly embracing virtualization and developing new features
to simplify virtualization techniques. First generation enhancements include Intel
Virtualization Technology (VT-x) and AMD's AMD-V which both target privileged 
instructions with a new CPU execution mode feature that allows the VMM to run in a privileged mode and thus  sensitive calls are set to automatically trap to the hypervisor, removing the need for either binary translation or paravirtualization\cite{website:vmware}

@@

Virtualization provides enables better IT infrastructure management with benefits like\cite{website:intel-VT}:

@@

Cloud Computing was a concept first publicly announced by John McCarthy in 1961 
(in a speech given to celebrate MIT's centennial). He saw a future in which computing 
power and even specific applications could be sold through the utility 
business model like water or electricity\cite{garfinkel1999architects}.
This concept and  supporting technologies evolved over the period and has taken us 
to the reality where computing resources are served as a utility. 
Technologies relating to CPU Power, network bandwidth, memory capacity 
and connectivity have improved dramatically in this period enabling Cloud Computing as 
a business model.

Cloud Computing is of interest among IT and network technologists
and academics, as well as CIOs and business leaders, due to its potential to 
reduce cost and risk, increase revenue, and enhance total customer experience.\cite{weinman2011future}
It has been difficult to give a concise definition for Cloud Computing 
and related terms. According to the U.S. National Institute of Standards 
and Technology (NIST),\cite{mell2009nist}  "Cloud computing is a model for enabling convenient,
on-demand network access to a shared pool of configurable computing resources
(e.g., networks, servers, storage, applications, and services) that can 
be rapidly provisioned and released with minimal management 
effort or service provider interaction."

@@

The essential characteristics are\cite{mell2009nist} :

@@

% need for hadoop

Demand for internet-scale\footnote{Petabytes of data in a distributed file system} analytics like search index building, result ranking, ad placement and personal recommendations have pushed for new platform architectures.

% what is hadoop

Apache Hadoop\cite{website:apache-hadoop} is a top-level Apache project that includes open source implementations of a distributed file system\cite{website:apache-hdfs} and MapReduce that was inspired by Google's GFS\cite{ghemawat2003google} and MapReduce\cite{dean2008mapreduce} projects.

@@

% users of hadoop
A wide variety of users have adopted Hadoop:

Facebook uses the Apache Hadoop platform to run it messages application\footnote {Facebook Messages 
integrates the display of all e-mail, SMS and chat messages between a pair or group of users.\cite{borthakur2011apache}.}. Hadoop was chosen because it met the application's reqirements for consistency, availability, partition tolerance, data model and scalability\cite{borthakur2011apache}.

Ebay uses the Apache Hadoop platform for data mining where they need to process petabytes of data to gain insights
from their transactinal data. Hadoop has helped EBay in changing how it collects, transforms, and uses data to generate business intelligence\cite {blog:ebay}.

The University of North Carolina at Chapel Hill uses Hadoop for Genomic Sequencing. 
The Hadoop system allows them to perform very custom analysis which is not found in a traditional business intelligence tool or in a SQL relational type of structure\cite{website:intel-hadoop}.

% benefits of hadoop

The advantage of this technology is that anybody can create an Hadoop cluster that
does fault tolerant and scalable distributed computing with off the shelf commodity hardware.

@@

Optimized for sequential reads of very large files (TB's - PB's)
Automatically manage data replication (3x by default) and recovery
Single view of the file system; any block can be accessed from any server

@@

% What is MapReduce
% Origin of MapReduce, why the need for MapReduce
% Implementations of MapReduce
% Apache Hadoop MapReduce architecture
% Applications that use MapReduce?
% MapReduce sucess stories / potential
